### 2018-12-09 Online Appendices/Appendix A.tex
Added all Box Plots

\documentclass[12pt, titlepage]{article}

\usepackage{fullpage} %full page typesetting
\usepackage{setspace} %allows for non-singlespacing
\usepackage{graphicx} %graphics capabilities
\usepackage{latexsym} %extra symbols
\usepackage{rotating} %rotation for figures
\usepackage{longtable} %tables that fill more than a single page
%\usepackage{hyperref} %hypertext links in the document
\usepackage{natbib} %better bibliographies
\usepackage{authblk} %author and affiliation in opening
\usepackage{mathpazo} %use palatino font, rather than times
\usepackage{appendix}
\usepackage{lscape}
\usepackage{tabulary}
\usepackage[nottoc]{tocbibind}
\usepackage[colorlinks=true,linkcolor=blue,citecolor=cyan]{hyperref}
\usepackage{ifthen}
\usepackage{float}
\usepackage{subcaption}
\usepackage{mwe}
\usepackage{caption}

\title{\tb{Place of Residence and Political Attitudes in Democracies Worldwide \\ {\large Online Appendix A -- Box Plots} }}

\author{Jennifer Lin}
\affil{New College of Florida}

\newcommand\e{\emph}
\newcommand\tb{\textbf}
\newcommand\un{\underline}
\newcommand\txt{\texttt}

\doublespacing

\begin{document}

\begin{singlespace}
\maketitle
\end{singlespace}

This Appendix features the box plots for each country divided by place of residence and by dependent variable.

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
	\centering 
	\includegraphics[width=\textwidth]{IdeoBP/Austria}
	\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
	\centering 
	\includegraphics[width=\textwidth]{BoxLib/Austria}
	\caption{Issue Stances}
	\end{subfigure}
	\caption{Austria}
	\label{Austria}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Bulgaria}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Bulgaria}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Bulgaria}
	\label{Bulgaria}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Czech}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Czech}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Czech Republic}
	\label{Czech}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Finland}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Finland}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Finland}
	\label{Finland}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/France}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/France}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{France}
	\label{France}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Germany}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Germany}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Germany}
	\label{Germany}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Greece}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Greece}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Greece}
	\label{Greece}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Iceland}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Iceland}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Iceland}
	\label{Iceland}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Israel}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Israel}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Israel}
	\label{Israel}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Japan}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Japan}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Japan}
	\label{Japan}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Latvia}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Latvia}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Latvia}
	\label{Latvia}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Mexico}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Mexico}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Mexico}
	\label{Mexico}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Montenegro}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Montenegro}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Montenegro}
	\label{Montenegro}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/NewZealand}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/NZealand}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{New Zealand}
	\label{NewZealand}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Peru}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Peru}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Peru}
	\label{Peru}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Philippines}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Philipines}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Philippines}
	\label{Philippines}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Poland}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Poland}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Poland}
	\label{Poland}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Romania}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Romania}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Romania}
	\label{Romania}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/SouthAfrica}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/SAfrica}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{South Africa}
	\label{SouthAfrica}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/SouthKorea}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/SKorea}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{South Korea}
	\label{SouthKorea}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Sweden}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Sweden}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Sweden}
	\label{Sweden}
\end{figure}

\begin{figure}[H]
	\centering
	\begin{subfigure}[b]{0.475\textwidth}   
		\centering 
		\includegraphics[width=\textwidth]{IdeoBP/Turkey}
		\caption{Self-Placement Ideology}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.475\textwidth}
		\centering 
		\includegraphics[width=\textwidth]{BoxLib/Turkey}
		\caption{Issue Stances}
	\end{subfigure}
	\caption{Turkey}
	\label{Turkey}
\end{figure}

\end{document}