### 2019-02-15 ModComparison/CompTableTEMPLATE.tex
TEMPLATE for All Polities table

\documentclass[12pt, titlepage]{article}

\usepackage{fullpage} %full page typesetting
\usepackage{setspace} %allows for non-singlespacing
\usepackage{graphicx} %graphics capabilities
\usepackage{latexsym} %extra symbols
\usepackage{rotating} %rotation for figures
\usepackage{longtable} %tables that fill more than a single page
\usepackage[colorlinks=true,linkcolor=blue,citecolor=cyan,urlcolor=red]{hyperref}
\usepackage{natbib} %better bibliographies
\usepackage{authblk} %author and affiliation in opening
\usepackage{mathpazo} %use palatino font, rather than times
\usepackage{tabulary}
\usepackage{float}
\usepackage{caption}

\title{\tb{Urban-Rural Divides, Democratic Satisfaction, and Citizen Attitudes Across Political Contexts \\ {\large CSES Inter-Module Comparisons} }}

\author{Jennifer Lin}
\affil{New College of Florida}

\newcommand\e{\emph}
\newcommand\tb{\textbf}
\newcommand\un{\underline}
\newcommand\txt{\texttt}

\begin{document}

\maketitle

\tableofcontents
\clearpage

%\listoftables
%\clearpage

\section{All Elections and Countries by Module}

\begin{center}
	\begin{longtable}{lcccc}
		\caption{\tb{All Elections and Countries by CSES Modules}\label{CSES-All} }\\
		\hline
		\tb{Country}&\tb{Module I}&\tb{Module II}&\tb{Module III}&\tb{Module IV}\\
		\hline
		\endfirsthead
		\multicolumn{4}{c}%
		{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
		\hline
		\tb{Country}&\tb{Module I}&\tb{Module II}&\tb{Module III}&\tb{Module IV}\\
		\hline
		\endhead
		\hline \multicolumn{4}{r}{\textit{Continued on next page}} \\
		\endfoot
		\hline
		\endlastfoot
		Albania&&&&\\
		Argentina&&&&\\
		Australia&&&&\\
		Belgium&&&&\\
		Belarus&&&&\\
		Brazil&&&&\\
		Bulgaria&&&&\\
		Canada&&&&\\
		Chile&&&&\\
		Croatia&&&&\\
		Czech Republic&&&&\\
		Denmark&&&&\\
		Estonia&&&&\\
		Finland&&&&\\
		France&&&&\\
		Germany&&&&\\
		Great Britain&&&&\\
		Greece&&&&\\
		Hong Kong&&&&\\
		Hungary&&&&\\
		Iceland&&&&\\
		Ireland&&&&\\
		Israel&&&&\\
		Italy&&&&\\
		Japan&&&&\\
		Kenya&&&&\\
		Kyrgyzstan&&&&\\
		Latvia&&&&\\
		Lithuania&&&&\\
		Mexico&&&&\\
		Montenegro&&&&\\
		New Zealand&&&&\\
		Netherlands&&&&\\
		Norway&&&&\\
		Peru&&&&\\
		Philippines&&&&\\
		Poland&&&&\\
		Portugal&&&&\\
		Romania&&&&\\
		Russia&&&&\\
		Serbia&&&&\\
		Slovakia&&&&\\
		Slovenia&&&&\\
		South Africa&&&&\\
		South Korea&&&&\\
		Spain&&&&\\
		Sweden&&&&\\
		Switzerland&&&&\\
		Taiwan&&&&\\
		Thailand&&&&\\
		Turkey&&&&\\
		Ukraine&&&&\\
		Uruguay&&&&\\
		United States&&&&\\
		\hline

	\end{longtable}
\end{center}

\end{document}